Making Sequential Programs Deterministic

### 2019-03-29 design.tex
ordernondet collection rules
todo: poly resolution rules

\begin{figure}
    $\infer[\rulename{iteration}]
    {\Gamma \vdash \|y| : \|OrderNonDet\ Iterator|\angles{\tau}}
    {\Gamma \vdash \|x| : \|OrderNonDet \ C\angles{\tau}|
    & C \sqsubseteq \|Collection|
    & \|y = C.iterator()|
    & \wellformed{\tau}
    }$

@@

    $\infer[\rulename{next element}]
    {\Gamma \vdash \|z| : \|NonDet \ \beta|}
    {\Gamma \vdash \|y| : \|OrderNonDet \ Iterator\angles{\kappa \ \beta}|
        & \|z = y.next()|
        & \wellformed{\kappa \ \beta}
    }$

@@

    $\infer[\rulename{hasnext element}]
    {\Gamma \vdash \|z| : \|Det \ boolean|}
    {\Gamma \vdash \|y| : \|OrderNonDet \ Iterator\angles{\tau}|
        & \|z = y.hasNext()|
        & \wellformed{\tau}
    }$

    $\infer[\rulename{size}]
    {\Gamma \vdash \|y| : \|Det \ int|}
    {\Gamma \vdash \|x| : \|OrderNonDet \ C\angles{\tau}|
        & C \sqsubseteq \|Collection|
        & \|y = x.size()|
        & \wellformed{\tau}
    }$
\caption{\<OrderNonDet> Collection rules}
\label{fig-ordernondet-rules}
\end{figure}

@@

\Cref{fig-poly-resolution} presents the rule for type checking of methods annotated with \<PolyDet> or its variants.
A polymorphic method type checks if every instantiation of \<PolyDet> type checks. If this method has \PolyDetUp\ or \PolyDetDown\
in its signature then the instantiation of \<PolyDet> at that location gets replaced by the appropriate type as shown in \cref{tab-poly-resolutions}.

Annotating \<size()> as \codeid{PolyDet$\downarrow$ int size(PolyDet List<T> list)} will only allow the following type resolutions:
\begin{itemize}
    \item \codeid{@Det int size(@Det List<T> list)}
    \item \codeid{@Det int size(@OrderNonDet List<T> list)}
    \item \codeid{@NonDet int size(@NonDet List<T> list)}
\end{itemize}

\begin{figure}
    $\infer[\rulename{poly resolution}]{\pi_r \ \beta_r\ \<m>(\overline{\pi_i \ \beta_i})\{\ \}}{\forall \kappa \in \{\|NonDet, OrderNonDet, Det|\} : \kappa \ \beta_r\ \<m>(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\|PolyDet|, \PolyDetUp, \PolyDetDown    \}}$
    \caption{Polymorphic resolution rule}
    \todo{I'm not sure what this is doing.  What code does it reject?}
    \todo{This suggests that \PolyDetUp is a qualifier, rather than
      $\uparrow$ being an operator on qualifiers.  We should clarify which
      it is.}
    \label{fig-poly-resolution}
\end{figure}

@@

\begin{figure}
    $\infer[\rulename{ordernondet polyup}]
    {\tau \uparrow \ = \|NonDet \ \beta|}
    {\tau \ = \|OrderNonNonDet \ \beta|
        & \wellformed{\tau}
    }$

    $\infer[\rulename{ordernondet polydown}]
    {\tau \downarrow \ = \|Det \ \beta|}
    {\tau \ = \|OrderNonNonDet \ \beta|
        & \wellformed{\tau}
    }$

@@

    $\infer[\rulename{polydown}]
    {\tau \downarrow \ = \kappa \ \beta}
    {\tau \ =\kappa \ \beta
        & \kappa \in \{\|Det, NonDet|\}
        & \wellformed{\tau}
    }$
\caption{Polymorphic resolution}
\label{tab-poly-resolutions}
\end{figure}

>>>>

\begin{figure}
%    $\infer[\rulename{iteration}]
%    {\Gamma \vdash \|y| : \|OrderNonDet\ Iterator|\angles{\tau}}
%    {\Gamma \vdash \|x| : \|OrderNonDet \ C\angles{\tau}|
%    & C \sqsubseteq \|Collection|
%    & \|y = C.iterator()|
%    & \wellformed{\tau}
%    }$

    \rulename{iteration}: 
    $
    \forall \tau . \|OrderNonDet \ C| \angles{\tau} \rightarrow 
    \|OrderNonDet \ Iterator| \angles{\tau}
    $

@@

%    $\infer[\rulename{next element}]
%    {\Gamma \vdash \|z| : \|NonDet \ \beta|}
%    {\Gamma \vdash \|y| : \|OrderNonDet \ Iterator\angles{\kappa \ \beta}|
%        & \|z = y.next()|
%        & \wellformed{\kappa \ \beta}
%    }$

    \rulename{next element}:
    $
    \forall \kappa \beta . \|OrderNonDet \ Iterator|\angles{\kappa \ \beta} \rightarrow
    \|NonDet|\ \beta
    $

@@

%    $\infer[\rulename{hasnext element}]
%    {\Gamma \vdash \|z| : \|Det \ boolean|}
%    {\Gamma \vdash \|y| : \|OrderNonDet \ Iterator\angles{\tau}|
%        & \|z = y.hasNext()|
%        & \wellformed{\tau}
%    }$

    \rulename{hasnext element}:
    $
    \forall \tau . \|OrderNonDet \ Iterator|\angles{\tau} \rightarrow
    \|Det \ boolean|
    $
    \bigskip

%    $\infer[\rulename{size}]
%    {\Gamma \vdash \|y| : \|Det \ int|}
%    {\Gamma \vdash \|x| : \|OrderNonDet \ C\angles{\tau}|
%        & C \sqsubseteq \|Collection|
%        & \|y = x.size()|
%        & \wellformed{\tau}
%    }$

    \rulename{size}:
    $
    \forall \tau . \|OrderNonDet \ C|\angles{\tau} \rightarrow
    \|Det \ int|
    $

\caption{\<OrderNonDet> Collection rules}
\label{fig-ordernondet-rules}
\end{figure}

@@

%\Cref{fig-poly-resolution} presents the rule for type checking of methods annotated with \<PolyDet> or its variants.
A polymorphic method type checks if every instantiation of \<PolyDet> type checks. If this method has \PolyDetUp\ or \PolyDetDown\
in its signature then the instantiation of \<PolyDet> at that location gets replaced by the appropriate type as shown in \cref{fig:poly-resolutions}.

Annotating \<size()> as \codeid{PolyDet$\downarrow$ int size(PolyDet List<T> list)} will only allow the following type instantiations:
\begin{itemize}
    \item \codeid{@Det int size(@Det List<T> list)}
    \item \codeid{@Det int size(@OrderNonDet List<T> list)}
    \item \codeid{@NonDet int size(@NonDet List<T> list)}
\end{itemize}

%\begin{figure}
%    $\infer[\rulename{poly resolution}]{\pi_r \ \beta_r\ \<m>(\overline{\pi_i \ \beta_i})\{\ \}}{\forall \kappa \in \{\|NonDet, OrderNonDet, Det|\} : \kappa \ \beta_r\ \<m>(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\|PolyDet|, \PolyDetUp, \PolyDetDown    \}}$
%    \caption{Polymorphic resolution rule}
%    \todo{I'm not sure what this is doing.  What code does it reject?}
%    \todo{This suggests that \PolyDetUp is a qualifier, rather than
%      $\uparrow$ being an operator on qualifiers.  We should clarify which
%      it is.}
%    \label{fig-poly-resolution}
%\end{figure}

@@

\begin{figure}
    $\infer[\rulename{ordernondet polyup}]
    {\tau \uparrow \ = \|NonDet \ \beta|}
    {\tau \ = \|OrderNonDet \ \beta|
        & \wellformed{\tau}
    }$

@@

    $\infer[\rulename{ordernondet polydown}]
    {\tau \downarrow \ = \|Det \ \beta|}
    {\tau \ = \|OrderNonDet \ \beta|
        & \wellformed{\tau}
    }$

    $\infer[\rulename{polydown}]
    {\tau \downarrow \ = \kappa \ \beta}
    {\tau \ =\kappa \ \beta
        & \kappa \in \{\|Det, NonDet|\}
        & \wellformed{\tau}
    }$
\caption{Polymorphic resolution ($\uparrow$ is overloaded for types and qualifiers)}
\label{fig:poly-resolutions}
\end{figure}
